Synthesizing Verified Components for Cyber Assured Systems Engineering \thanks{{DISTRIBUTION STATEMENT A. Approved for public release.} }

### 2021-05-11 paper/abstract.tex
Fix typos
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they recover or safely shut down as requirements dictate. The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in this cyber-resiliency in a Model-Based Systems Engineering environment. This paper discusses the automated model transformations that add in high-assurance components. These components are filters and monitors that prevent malicious input and detect supply chain attacks respectively. A formal specification defines each component and is used to verify that the component addresses system level cyber-requirements. Implementations are synthesized from the specification, and that synthesis is proven to preserve the exact meaning of the specification. The transformations are integrated into the Open Source AADL Tool Environment (OSATE). The paper further reports a case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning. Here the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

>>>>

Cyber-physical systems, such as avionics, must be tolerant to cyber-attacks in the same way they are tolerant to random faults: they either recover or safely shut down as requirements dictate. The DARPA Cyber Assured Systems Engineering program is developing design, analysis, and verification tools that enable systems engineers to design-in cyber-resiliency in a Model-Based Systems Engineering environment. This paper describes automated model transformations that introduce  high-assurance cyber-resiliency components into a system, in particular filters and monitors that prevent malicious input and detect supply chain attacks, respectively. A formal specification defines each component and is used to verify that the component addresses system level cyber requirements. Implementations are synthesized from the specifications, and that synthesis is proven to preserve the exact meaning of the specifications. The transformations are integrated into the Open Source AADL Tool Environment (OSATE). The paper further reports a case study applying these transformations to a UAV system that uses the Air Force Research Laboratory's OpenUxAS services for route planning. In the case study, the transforms add filters to guard against malformed input and monitors to guard against ground station spoofing and malicious flight plans from OpenUxAS.

### 2021-05-11 paper/contig.tex
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types} to define the input and output data (cite contiguity). A contiguity type is a self-describing specification for messages. Its formalism has basis in formal languages. Similar to how a regular expression implies a set of words that form its language, so does a contiguity type specification imply a set of messages for its language where a message is a finite sequence of contiguous bytes (e.g., a string).

>>>>

The formal specification of a component, and the synthesis of that specification, relies on \emph{contiguity types}~\cite{contiguity-types} to define the input and output data (cite contiguity). A contiguity type is a self-describing specification for messages. Its formalism has basis in formal languages. Similar to how a regular expression implies a set of words that form its language, so does a contiguity type specification imply a set of messages for its language where a message is a finite sequence of contiguous bytes (e.g., a string).

### 2021-05-11 paper/example.tex
Fix fig and add link to repo
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

\figref{fig:example} is an diagram form an AADL model of an implementation of a software system  (SW) for route planning and automated control for a UAV. It is loosely based on the system in the case study introduced later. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component.

@@

There are two requirements added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request. Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.

AGREE model checks the composition of the new specifications for the implementation in \figref{fig:example}(a) to see if it complies with the added cyber-resilience properties. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager. The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request. A clear malicious behavior from the untrusted component.

>>>>

\figref{fig:example} is an AADL model of an implementation of a software system (SW) for route planning and automated control for a UAV. It is loosely based on the system in the case study introduced later.  The source for the entire model is found at \cite{repo}. The system receives an automation request that is forwarded to an untrusted third-party route planner (AI) that decides the flight path of the UAV based on its current position and the requested task. The waypoint manager (WM) receives the mission command as a set of waypoints from the planner and starts the UAV flying the mission, issuing waypoints to the UAV flight controller as the UAV location changes. It is an \emph{as is} legacy component.

@@

Two requirements are added to the SW AGREE specification related to the cyber-vulnerability analysis: the first, \emph{Waypoint is well-formed} disallows malformed waypoints from being propagated downstream; and the second, \emph{Alert if start is not bounded relative to a request}, disallows a mission being started without a request or if the start is delayed more than one step after a request. Here the goal is to make clear that a delayed or unrequested automation response may prevent the UAV from flying its intended mission or to fly the wrong mission.

AGREE examines the composition of the new specifications for the implementation in \figref{fig:example}(a) to determine whether it complies with the added cyber-resilience properties, by way of model checking. The results of the model checker are in \figref{fig:example}(b). The red exclamation points are properties that do not hold. Each comes with a counter-example. The results are not unexpected given the AI route planner's unconstrained behavior and the new assumptions about input to the legacy waypoint manager. The counter-example for \emph{Alert if start is not bounded relative to a request} shows a response in the first time step with no matching request. A clear malicious behavior from the untrusted component.

### 2021-05-11 paper/intro.tex
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems. 
The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.  
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.  Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

>>>>

The DARPA Cyber Assured Systems Engineering (CASE) program is targeted to developing design, analysis, and verification tools that enable systems engineers to {\em design-in} cyber-resiliency for complex cyber-physical systems.\footnote{The
  views expressed are those of the authors and do not reflect the
  official policy or position of the Defense Advanced
  Research Projects Agency (DARPA) or the U.S. Government.}
The result is a Model-Based Systems Engineering (MBSE) environment called {\em BriefCASE} which is based on the Architecture Analysis and Design Language (AADL)~\cite{aadl}.  BriefCASE extends the Open Source AADL Tool Environment (OSATE) to add new design, analysis, and code generation capabilities targeted at building cyber-resilient systems.

BriefCASE provides access to two analysis tools (GearCASE~\cite{gearcase2020} and DCRYPPS~\cite{dcrypps2019}) that can examine AADL models to detect potential cyber vulnerabilities and suggest requirements for mitigation.  
A library of architectural transforms guides systems engineers through automated model transformations that modify the architecture to address these requirements, possibly inserting new high-assurance components into the system. 
Implementations for these new high-assurance components are synthesized from formal specifications using the Semantic Properties for Language and Automata Theory (SPLAT) tool~\cite{slind-hcss2020},~\cite{formal-filter-synth-langsec}.
Formal verification that the transformed system model satisfies its cyber requirements is accomplished via model checking using the Assume Guarantee Reasoning Environment (AGREE)~\cite{agree2013}.  Cyber-resilient code implementing the verified model is automatically generated using the High Assurance Modeling and Rapid Engineering for Embedded Systems (HAMR) toolkit~\cite{hamr}.  If desired, this code can be targeted to the formally verified seL4 secure microkernel~\cite{sel4-2009}.

### 2021-05-11 paper/paper.tex
Add contig types
Restructure tex files
Update with Hardin feedback
DARPA Public Release boilerplate, author reordering, bib additions, misc edits

\begin{document}
\title{
  Synthesizing Verified Components for Cyber Assured Systems Engineering 
  \thanks{
    Funded by DARPA Cyber Assured Systems Engineering (CASE) contract HR00111890001.
  }
}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{
Konrad Slind, 
Junaid Babar, 
Isaac Amundson, and
David Hardin}
\IEEEauthorblockA{
Trusted Systems Group   \\
Collins Aerospace       \\
Minneapolis, Minnesota  \\
}
\and
\IEEEauthorblockN{Eric Mercer}
\IEEEauthorblockA{Department of Computer Science\\
Brigham Young University\\
Provo, Utah}
}

% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

@@

% \section{Introduction}
% \input{intro}

@@

% \section{Contiguity Types}
% \input{contig}

% \section{Filters}
% \input{filters}

% \section{Monitors}
% \input{monitors}

% \section{OSATE Integration for AADL}
% \input{aadl-agree-osate}

@@

\section*{Acknowledgments}
This work is funded by DARPA contract HR00111890001. The views, opinions and/or findings expressed are those of the authors and
should not be interpreted as representing the official views or policies of the Department of Defense or the U.S. Government.

@@

% references section
\bibliographystyle{IEEEtran}
\bibliography{paper}
\end{document}

>>>>

\begin{document}
\title{
  Synthesizing Verified Components for Cyber Assured Systems Engineering 
  \thanks{{DISTRIBUTION STATEMENT A.  Approved for public release.}
  }
}

% author names and affiliations
% use a multiple column layout for up to three different
% affiliations
\author{
\IEEEauthorblockN{Eric Mercer}
\IEEEauthorblockA{Department of Computer Science\\
Brigham Young University\\
Provo, Utah}
\and
\IEEEauthorblockN{
Isaac Amundson and
Konrad Slind}
\IEEEauthorblockA{
Trusted Systems Group   \\
Collins Aerospace       \\
Minneapolis, Minnesota}
\and
\IEEEauthorblockN{
Junaid Babar and 
David Hardin}
\IEEEauthorblockA{
Trusted Systems Group   \\
Collins Aerospace       \\
Cedar Rapids, Iowa}
}
% conference papers do not typically use \thanks and this command
% is locked out in conference mode. If really needed, such as for
% the acknowledgment of grants, issue a \IEEEoverridecommandlockouts
% after \documentclass

@@

\section{Introduction}
\input{intro}

@@

\section{AGREE Specification}
\input{agree}

\section{Synthesis}
\input{synthesis}

\section{UxAS Case Study}
\input{case-study}

@@

\section*{Acknowledgments}
This work was sponsored in part by the Defense Advanced Research
Projects Agency (DARPA).

@@

% references section
\bibliographystyle{IEEEtran}
\bibliography{paper}

\appendix
\section{Contiguity Types}
\input{contig}

\end{document}
